Constraints on tensor and scalar couplings from $B\to K\bar\mu\mu$ and $B_s\to \bar\mu\mu$

### 2015-07-09 tensor/draft/draft-tensor-op.tex
[tensor] Complete review of section IV

In the following we consider a model-independent scenario as predicted by the SM
augmented by dimension six operators that respect the SM gauge group and are
composed out of SM fields only. After the discovery of a scalar resonance at the LHC
that complies currently with all requirements of the Higgs particle in the SM
and the steadily rising lower bounds on new particle masses towards higher energies
from ATLAS and CMS such a scenario becomes more and more viable.

A non-redundant set of dimension six operators of
this effective theory (SM-EFT) was given in \cite{Grzadkowski:2010es} and the
matching to the effective theory of $\Delta B = 1$ decays \refeq{eq:Heff} at the
scale $\mu \sim m_W$ of the order of the $W$-boson mass in~\cite{D'Ambrosio:2002ex}
for vector couplings $\wilson{7,\,9,\,10}$. The matching for tensor
\refeq{eq:tensor:ops} and scalar \refeq{eq:scalar:ops} operators
\cite{Alonso:2014csa} shows that SM gauge groups in conjunction with the linear
representation impose the relations
\begin{align}
  \label{eq:mEFT-constraints}
  \wilson{P}  & = -\wilson{S}\,, &
  \wilson{P'} & = \wilson{S'}\,, &
  \wilson{T}  & = \wilson{T5} = 0
\end{align}
on scalar and pseudo-scalar couplings and require tensor couplings to be
suppressed to the level of dimension eight operators. In consequence only two
scalar couplings $\wilson{S,\, S'}$ arise that scale as $\sim v^2/\Lambda_{\rm NP} \ll 1$
where $v^2$ denotes the scale of electroweak symmetry breaking and $\Lambda_{\rm NP}$
a high energy scale at which new physics generates these operators.

@@

The SM gauge groups do not impose constraints on the operators of the SM-EFT that
imply relations among $\wilson{10}$ and $\wilson{10'}$, so they
are in general additional independent parameters. Here we find that destructive
interference with contributions involving $\wilson{10,\, 10'}$ do not change
much upper bounds on $|\wilson{S,\, S'}|$ found in \cite{Alonso:2014csa}. Since we
consider here complex-valued couplings the allowed regions are circles rather
than rings as for the case of real-valued couplings. The result of the simultaneous
fit of complex-valued $\wilson{S,\,S',\,10,\,10'}$ is shown in
\reffig{fig:scalar:SM-EFT}.

@@

The most general fit of complex-valued tensor and scalar couplings
$\wilson{S,S',P,P',T,T5}$ in combination with complex-valued vector couplings
$\wilson{9,\,9',\,10,\, 10'}$ --- i.e. the combination of \refsec{sec:tensor}
and \refsec{sec:scalar} with same prior ranges --- does not yield very different
bounds. The changes are only small and in fact the bounds tend to be even more
stringent, because tensor and scalar couplings can contribute to $F_H^{\mu}$
only constructively --- see \refeq{eq:FH-dependence}.
As before branching-ratio measurements of $B\to K^{*}\bar\mu\mu$ help to strengthen
the constraints on $\wilson{T,\, T5}$. This demonstrates that even in the case
of complex-valued couplings the data is able to constrain all real and imaginary
parts.

@@

Eventually, we present the yet allowed ranges due to non-standard tensor
and scalar couplings of the so far unmeasured angular observables $J_{6c}$,
$(J_{1s} - 3\, J_{2s})$ and $(J_{1c} + J_{2c})$ in $B\to K^*(\to K\pi)\bar\ell\ell$
in order to explore the potential of their future measurements in fits of the kind
pursued in this work.
The 68\% probability intervals are collected in \reftab{tab:postdictions}
for the various model-independent scenarios, allowing for interference with
non-zero vector couplings, in \refsec{sec:tensor}, \refsec{sec:scalar}
and \refsec{sec:tensor:scalar} for $q^2$-integrated quantities
in the low- and high-$q^2$ bins $q^2\in [1.1,\, 6]$ GeV$^2$ and $q^2\in [15,\, 19]$
GeV$^2$, respectively. Their posterior distributions have been obtained from the
final posterior samples as explained in~\cite{Beaujean:2012uj}.

For comparison, the SM estimates are also given in \reftab{tab:postdictions}
multiplied by the $B^0$-meson life time $\tau_{B^0} = 1.519$~ps.
At low-$q^2$ they are of order ${\cal O}(10^{-9})$ for $(J_{1s} - 3\, J_{2s})$ and
$(J_{1c} + J_{2c})$, getting suppressed down to ${\cal O}(10^{-10})$ at high-$q^2$
due to the $m_\ell/\sqrt{q^2}$ dependence. In order to compare it to the typical
magnitude of vector couplings, note that for example the branching ratio of
$B\to K^*\bar\ell\ell$
\begin{align}
  \frac{{\rm d}\BR}{{\rm d}q^2} & = \frac{\tau_{B^0}}{3} \left[
    2\, (3 J_{1s} - J_{2s}) + (3 J_{1c} -  J_{2c})
  \right]
\end{align}
is numerically of the order ${\cal O}(10^{-7})$ in the SM
\begin{equation}
\begin{aligned}
  \langle {\rm d}\BR/{\rm d}q^2 \rangle_{[1.1,\, 6]} & =
  \left(1.86_{-0.44}^{+0.53} \right) \cdot 10^{-7} \,,
\\
  \langle {\rm d}\BR/{\rm d}q^2 \rangle_{[15,\, 19]} & =
  \left(1.91_{-0.16}^{+0.16} \right) \cdot 10^{-7} \,,
\end{aligned}
\end{equation}
when integrated over low- and high-$q^2$ regions. \todo{Elaborate on larger
theory uncertainties at low-$q^2$, due to subleading corrections.
Form factor uncertainties are small due to correlations, would be larger
when neglecting them.}

The angular observable $J_{6c}$
is strictly zero for the approximation adopted here. Non-zero contributions
can be generated by QED corrections or potentially from higher-dimensional
($d \geq 8$) $|\Delta B| = |\Delta S| = 1$ operators, leading to parametrical
suppression of $\alpha_e/(4 \pi)$ or $(m_b/m_W)^2$, respectively. Such a
suppression must be compared to potential suppression present for tensor
and scalar contributions in particular NP models in order to decide their
relevance. Our model-independent fits are still in a realm where such
considerations are insignificant since current experimental measurements, in
combination with theory uncertainties, do not impose yet too stringent
constraints on tensor and scalar couplings.

Beyond the SM, $J_{6c}$ can become of order ${\cal O}(10^{-9})$ in scenarios
involving tensor couplings and only ${\cal O}(10^{-10})$ in their absence.
Concerning $(J_{1c} + J_{2c})$ and $(J_{1s} - 3\, J_{2s})$, again substantial
deviations from the SM prediction are possible only in the presence of non-zero
tensor couplings. In this case enhancement by two orders of magnitude are
possible up to the ${\cal O}(10^{-8})$ at high-$q^2$ and also at low-$q^2$ in
the case of $(J_{1s} - 3\, J_{2s})$, always being consistent with all included
experimental constraints in \reftab{tab:observables}. In view of current
experimental precision of $20\%$ of the branching ratio measurements at LHCb
\cite{Aaij:2013iag} with only 1 fb$^{-1}$, corresponding to the ${\cal O}(10^{-8})$, one can
indeed hope for some sensitivity to such large effects in $(J_{1c} + J_{2c})$
and $(J_{1s} - 3\, J_{2s})$ for the not yet published 3 fb$^{-1}$ data set.
\todo{Would be a nice place to cite method of moments paper?}

>>>>

In the following we consider a model-independent scenario as predicted by the SM
augmented by dimension six operators that respect the SM gauge group and are
composed out of SM fields only. After the discovery of a scalar resonance at the LHC
that complies currently with all requirements of the Higgs particle in the SM
and the steadily rising lower bounds on new particle masses towards higher energies
from ATLAS and CMS such a scenario becomes more and more viable \fred{hard to read}.

A non-redundant set of dimension-six operators of this effective
theory (SM-EFT) was given in \cite{Grzadkowski:2010es} and the
matching to the effective theory of $\Delta B = 1$ decays
\refeq{eq:Heff} at the scale $\mu \sim m_W$ of the order of the
$W$-boson mass in~\cite{D'Ambrosio:2002ex} for vector couplings
$\wilson{7,\,9,\,10}$. The matching for tensor \refeq{eq:tensor:ops}
and scalar \refeq{eq:scalar:ops} operators \cite{Alonso:2014csa} shows
that SM gauge groups in conjunction with the linear representation
\todo{Ask Martin, Oscar how to correctly phrase this} impose the
relations
\begin{align}
  \label{eq:mEFT-constraints}
  \wilson{P}  & = -\wilson{S}\,, &
  \wilson{P'} & = \wilson{S'}\,, &
  \wilson{T}  & = \wilson{T5} = 0
\end{align}
on scalar and pseudo-scalar couplings and require tensor couplings to be
suppressed to the level of dimension eight operators. In consequence only two
scalar couplings $\wilson{S,\, S'}$ arise that scale as $\sim v^2/\Lambda_{\rm NP} \ll 1$
where $v^2$ denotes the scale of electroweak symmetry breaking and $\Lambda_{\rm NP}$
a high energy scale at which new physics generates these operators.

@@

The SM gauge group does not imply any relation between $\wilson{10}$
and $\wilson{10'}$, so they are in general additional independent
parameters. Here we find that destructive interference with
contributions involving $\wilson{10,\, 10'}$ does not significantly
alter the bounds on $\wilson{S,\, S'}$.  Since we consider here
complex-valued couplings the allowed regions are circles rather than
rings as for the case of real-valued
couplings~\cite{Alonso:2014csa}. The result of the simultaneous fit of
complex-valued $\wilson{S,\,S',\,10,\,10'}$ is shown in
\reffig{fig:scalar:SM-EFT}. \fred{Circle smaller, prob. moves to
  center compared to Alonso}

@@

The most general fit of complex-valued tensor and scalar couplings
$\wilson{S,S',P,P',T,T5}$ in combination with complex-valued vector
couplings $\wilson{9,\,9',\,10,\, 10'}$ --- i.e. the combination of
\refsec{sec:tensor} and \refsec{sec:scalar} with same prior ranges ---
yields bounds very similar to those in tables \ref{tab:cTT5:1-dimCLs}
and \ref{tab:cSP:1-dimCLs} . The changes are only small and in fact
the bounds tend to be even more stringent because tensor and scalar
couplings can contribute to $F_H^{\mu}$ only constructively --- see
\refeq{eq:FH-dependence}.  As before, branching-ratio measurements of
$B\to K^{*}\bar\mu\mu$ help to strengthen the constraints on
$\wilson{T,\, T5}$. This demonstrates that even in the case of
complex-valued couplings the data are able to constrain all real and
imaginary parts.

@@

Now we present the yet allowed ranges due to non-standard tensor and
scalar couplings of the so far unmeasured angular observables
$J_{6c}$, $(J_{1s} - 3\, J_{2s})$ and $(J_{1c} + J_{2c})$ in $B\to
K^*(\to K\pi)\bar\ell\ell$ in order to explore the potential of their
future measurements in fits of the kind pursued in this work.  The
68\% probability intervals are collected in \reftab{tab:postdictions}
for the various model-independent scenarios, allowing for interference
with non-zero vector couplings, in \refsec{sec:tensor},
\refsec{sec:scalar} and \refsec{sec:tensor:scalar} for
$q^2$-integrated quantities in the low- and high-$q^2$ bins $q^2\in
[1.1,\, 6]$ GeV$^2$ and $q^2\in [15,\, 19]$ GeV$^2$,
respectively. Their posterior distributions have been obtained from
the final posterior samples as explained in~\cite{Beaujean:2012uj}
\todo{check uncertainty propagation there}.

For comparison, the SM estimates are also given in \reftab{tab:postdictions}
multiplied by the $B^0$-meson life time $\tau_{B^0} = 1.519$~ps.
At low-$q^2$ they are of order ${\cal O}(10^{-9})$ for $(J_{1s} - 3\, J_{2s})$ and
$(J_{1c} + J_{2c})$, getting suppressed down to ${\cal O}(10^{-10})$ at high-$q^2$
due to the $m_\ell/\sqrt{q^2}$ dependence. In order to compare it to the typical
magnitude of vector couplings, note that for example the branching ratio of
$B\to K^*\bar\ell\ell$
\begin{align}
  \frac{{\rm d}\BR}{{\rm d}q^2} & = \frac{\tau_{B^0}}{3} \left[
    2\, (3 J_{1s} - J_{2s}) + (3 J_{1c} -  J_{2c})
  \right]
\end{align}
is approximately equal to $2 \cdot 10^{-7}$ in the SM for both the
$q^2 \in [1,\, 6]$ and $[15,\, 19]$ GeV$^2$ bins.

The angular observable $J_{6c}$ is strictly zero in the absence of
tensor and scalar couplings. Non-zero contributions can be generated
in the SM by QED corrections or potentially from higher-dimensional
($d \geq 8$) $|\Delta B| = |\Delta S| = 1$ operators, leading to
parametrical suppression of $\alpha_e/(4 \pi)$ or $(m_b/m_W)^2$,
respectively. Such a suppression must be compared to potential
suppression present for tensor and scalar contributions in particular
NP models in order to decide their relevance. Our model-independent
fits are still in a realm where such considerations are insignificant
since current experimental measurements, in combination with theory
uncertainties, do not impose yet too stringent constraints on tensor
and scalar couplings.

Beyond the SM, $J_{6c}$ can become of order ${\cal O}(10^{-9})$ in
scenarios involving tensor couplings only and about ${\cal
  O}(10^{-10})$ in the presence of scalar couplings only. Both effects
are due to interference with vector couplings. Due to the strong
suppression of scalar couplings, the direct interference of scalar and
tensor contributions the allowed region in the most general scenario
\todo{schematic formula from notes}

Concerning $(J_{1c} + J_{2c})$ and $(J_{1s} - 3\, J_{2s})$, again
substantial deviations from the SM prediction are possible only in the
presence of non-zero tensor couplings. In this case enhancement by two
orders of magnitude are possible up to the ${\cal O}(10^{-8})$ at
high-$q^2$ and also at low-$q^2$ in the case of $(J_{1s} - 3\,
J_{2s})$, always being consistent with all included experimental
constraints in \reftab{tab:observables}. In view of current
experimental precision of $20\%$ of the branching ratio measurements
at LHCb \cite{Aaij:2013iag} with only 1 fb$^{-1}$, corresponding to
the ${\cal O}(10^{-8})$, one can indeed hope for some sensitivity to
such large effects in $(J_{1c} + J_{2c})$ and $(J_{1s} - 3\, J_{2s})$
for the not yet published 3 fb$^{-1}$ data set.  \todo{Would be a nice
  place to cite method of moments paper?}

The CP asymmetry $A_{\Delta\Gamma}(B_s\to \bar\mu\mu)$ induced by
non-vanishing width of the $B_s$ meson (cf. \refapp{app:BKstarellell})
can still take any value in the range $[-1,\, 1]$ in scenarios with
nonzero scalar couplings.
